Fast, Scalable Phrase-Based SMT Decoding

### 2016-05-04 acl.2016/acl2016.tex
changes after review

Firstly, the search algorithm creates and destroy a large number of hypothesis and related objects. The memory allocation required puts a burden on the operating system due to the need to synchronize  memory access. This problem worsens with multi-threaded applications such as Moses.

@@

Secondly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.

@@

Thirdly, we will look at optimizing the phrase-table.

@@

We propose two optimizations. One, the default translation rule caching mechanism in Moses saves the most recently used rules. However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}

@@

Two, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule. Compression was championed by ~\newcite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads. We shall therefore take the opposite approach to and explore optimizing decoding speed by disabling compression.

@@

As a baseline, we use the latest version of the Moses decoder taken from the github repository.

For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz). The server has 380GB RAM. The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library\footnote{http://goog-perftools.sourceforge.net/doc/tcmalloc.html}.

@@

We create a dynamic memory pool which can grow as more memory is requested. The memory is not released, instead the pools are reset and re-used. We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding of each sentence. Data structures are created in either pool according to their life cycle.

For critical data structures with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used. This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.

Over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, Table~\ref{tab:optimizing-memory}, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}[h]
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, is coverage cardinality, ie. hypotheses that have translated the same number of source words are stored in the same stack. There have been research into other stack layouts such as~\cite{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder in~\cite{Brown:1993} uses coverage stacks, as opposed to coverage cardinality.

>>>>

The search algorithm creates and destroy a large number of hypothesis and related objects. The memory allocation required puts a burden on the operating system due to the need to synchronize  memory access. This problem worsens with multi-threaded applications such as Moses.

@@

The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, is coverage cardinality, that is, hypotheses that have translated the same number of source words are stored in the same stack. Since stack pruning occurs in each stack independently of each other, this affect the decoder's search errors.

However, there are alternatives to the coverage cardinality configuration. ~\newcite{Och:2001a} uses a single stack for all hypotheses, ~\newcite{Brown:1993} uses coverage stacks (ie. one stack per unique coverage vector) while ~\newcite{peitz2012jane} and ~\newcite{Zens+Ney:2008:iwslt} apply both coverage and cardinality pruning. While useful, these prior works present only one particular stack configuration each.

~\newcite{ortizmartinez-garciavarea-casacuberta:2006:WMT} explore a range of stack configurations by defining a granularity parameter which controls the maximum number of stacks required to decode a sentence.

We shall re-visit the question of stack configuration with a particular emphasis on how they can help improve the tradeoff between speed and translation quality. We will do so in the context of the cube-pruning algorithm, the algorithm that we will be using and which was not available to many of the earlier work.

@@

@@

We propose two optimizations. Firstly, the default translation rule caching mechanism in Moses saves the most recently used rules. However, this require locking and periodic clearing of old rules in order to reduce memory consumption resulting in \emph{slower} decoding, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}

@@

Secondly, the Probing phrase-table use a simple compression algorithm to compress the target side of the translation rule. Compression was championed by ~\newcite{junczys_tsd_2012b} as the main reason behind the speed of their phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads. We shall therefore take the opposite approach to and explore optimizing decoding speed by disabling compression.

@@

As a baseline, we use a recent\footnote{between January and May 2016} version of the Moses decoder taken from the github repository.

For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz). The server has 380GB RAM. The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost 1.59\footnote{http://boost.org/} and the tcmalloc library.

@@

We instantiate two memory pools per decoding thread, one which is never reset and another which is reset after the decoding of each sentence. Data structures are created in either pool according to their life cycle.

Hypothesis objects and the data structures it contains are re-used by maintaining a thread-specific LIFO queue of unused, recombined and pruned objects from which new hypotheses are allocated in preference to create new hypotheses in the memory pool.

Each sentence is decoded from start to finish using one worker thread. The CPU affinity of each thread is set to a specific core to minimize memory page faults as a result of threads switching cores.

As can be seen in Table~\ref{tab:optimizing-memory}, over 24\% of the Moses decoder running time is spent on memory management and this increases to 39\% when 32 threads are used, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}[h]
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@
